---
title: "Intro to R Packages"
output:
  html_document:
    df_print: paged
---

## Installing & Exploring Packages - seqinr

This notebook was created based on instructions at https://bcb420-2022.github.io/R_basics/r-install.html#task-5---experiment-with-rstudio-and-packages.

### Installing & Loading Packages

```{r}
library()
```

By running the command above and examining the page that it opens, I see that seqinr is not installed.

<br />

```{r}
search()
```

This command showed that I have ten packages loaded.

<br />

```{r}
?vignette

```

This opened a help page in the 'Help' tab that describes how to view and list vignettes.

<br />

```{r}
?install.packages
if (! requireNamespace("seqinr", quietly=TRUE)) {
  install.packages("seqinr",repos="https://cran.rstudio.com/")
}
```

The above code opened a help page which described how to install new packages. The second command successfully installed seqinr (and will only install it if it is not already installed).

<br />

### Listing Functions and Datasets

```{r}
library(help="seqinr") 

library(seqinr)
ls("package:seqinr")
```

The top command opened the vignette for seqinr in a new page. 
The bottom two commands allowed me to list the 203 functions available in seqinr.

<br />

```{r}
?seqinr::a 
seqinr::a("Tyr")
```

The above commands opened help documentation for the `a` method in `seqinr`. 
This method converts AA 3-letter codes to 1-letter codes (ex. Tyr -> Y).

<br />

```{r}
seqinr::words(3, c("A", "G", "C", "U"))
```

The above code produced a vector with all possible 3-letter combinations of 
"A", "G", "C", and "U". In other words, it produced all possible RNA codons.

<br />

```{r}
?data
```

According to the help page, the `data` function loads or lists available data sets in a package.

<br />

```{r}
data(package="seqinr")
```

One dataset available in the `seqinr` package is ec999, with has "999 coding sequences from E. coli".

<br />

```{r}
data(aaindex, package="seqinr")   
```

The aaindex dataset lists 544 physicochemical and biological properties for the 20 AAs (ex. conformation parameter of inner helix, average volume of buried residue, melting point, size, etc.).

<br />

```{r}
?aaindex
```

The above command opens the help documentation for the `aaindex` dataset.

<br />

```{r}
aaindex$FASG890101
aaindex$PONJ960101 
```

FASG890101 and PONJ960101 are the accession numbers of two properties (of the 544) in the `aaindex` db. The first property describes hydrophobicity of the AAs, and the second property describes the average volumes of the AAs.

<br />

```{r}
plot(aaindex$FASG890101$I,
     aaindex$PONJ960101$I,
     xlab="hydrophobicity", ylab="volume", type="n")
text(aaindex$FASG890101$I,
     aaindex$PONJ960101$I,
     labels=a(names(aaindex$FASG890101$I)))
```

This command plots the hydrophobicity vs. the volume of each AA.

<br />

```{r}
seqinr::choosebank("swissprot")
mySeq <- seqinr::query("mySeq", "N=MBP1_YEAST")
mbp1 <- seqinr::getSequence(mySeq)
seqinr::closebank()
x <- seqinr::AAstat(mbp1[[1]])
barplot(sort(x$Compo), cex.names = 0.6)
```

This code block fetches the MBP1_YEAST sequence from the swissprot databank, and then plots some statistics about the AAs found in the sequence.

<br />

## Finding Packages -- sos

```{r}
if (! requireNamespace("sos", quietly=TRUE)) {
    install.packages("sos")
}
library(help = sos)       # basic information
browseVignettes("sos")    # available vignettes
 
sos::findFn("moving average")
```

The code above installed the `sos` package and listed 300 packages related to "moving average".

Other resources for finding packages are the CRAN Task View, Bioconductor Views, Metacran, MRAN, and Google.

<br />

## Self-Evaluation

1. The purpose of wrapping an `install.packages` command inside a `requireNamespace` conditional block is to ensure that the package is only installed if it is not already installed. This prevents RStudio from time-consuming and unnecessary installations.

<br />
<br />
