---
title: "Assignment 1"
author: Evgeniya Gorobets
date: "`r format(Sys.time(), '%d %b %Y')`"
output:
  html_document:
    df_print: paged
---

## GEOmetadb Setup

### Installing GEOmetadb

The following code is based off slide 7 in Professor Ruth Isserlin's Lecture 3
slides ("Finding Expression Data") (Isserlin, 2022). This is just to ensure that
`GEOmetadb`is installed and available whenever I want to run this notebook.
```{r}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
  install.packages("BiocManager")
}
if (!requireNamespace("GEOmetadb", quietly = TRUE)) {
    BiocManager::install("GEOmetadb")
}
```

### Connecting to the Database

TODO: add references to this code
```{r}
# Change the working directory if needed
if (getwd() == "/home/rstudio") {
  setwd("./projects/Assignment1")
}

sqlFilePath <- "data/GEOmetadb.sqlite"
if (!file.exists(sqlFilePath)) {
  options(timeout=600)
  GEOmetadb::getSQLiteFile(destdir="data")
}
con <- DBI::dbConnect(RSQLite::SQLite(), sqlFilePath)
```


## References

Ruth Isserlin's Lecture 3
slides ("Finding Expression Data") (Isserlin, 2022)

