---
title: "Assignment 1: Dataset Selection and Initial Processing"
author: Evgeniya Gorobets
date: "`r format(Sys.time(), '%d %b %Y')`"
output:
  html_document:
    df_print: paged
---

## Preface

This R Notebook was developed by Evgeniya Gorobets as part of an assessment for 
2022 BCB420H: Computational Systems Biology, University of Toronto, Toronto, CA.
Specifically, this notebook was the final submitted product for Assignment 1 
of the course.

This notebook only includes the download, cleaning, and processing of my 
selected GEO dataset. To view the code used to explore GEO, evaluate various 
GSEs, and ultimately select this dataset, see the R Notebook located at 
https://github.com/bcb420-2022/Evgeniya_Gorobets/blob/main/Assignment1/ExploringGEO.Rmd.

## Setup

The code below was taken from the R Notebook I did as a pre-requisite to this 
notebook, titled "Choosing a GEO Dataset". You can view the original code 
[here](https://github.com/bcb420-2022/Evgeniya_Gorobets/blob/main/Assignment1/ExploringGEO.Rmd).

```{r}
# Install BiocManage and GEOquery if required
if (!requireNamespace("BiocManager", quietly = TRUE)) {
  install.packages("BiocManager")
}
if (!requireNamespace("GEOquery", quietly = TRUE)) {
    BiocManager::install("GEOquery")
}

# Change working directory, if needed
if (getwd() == "/home/rstudio") {
  setwd("./projects/Assignment1")
}

# Set constants
GSE <- 'GSE129943'
DATA_DIR <- 'data'
```


## Downloading the Data

Add some commentary here

```{r}
gseDir <- paste0(DATA_DIR, "/", GSE)

# Download the supplemental files for GSE129943 if they're not yet in data dir
if (!file.exists(gseDir)) {
  dataFiles <- GEOquery::getGEOSuppFiles(GSE, baseDir=DATA_DIR)
  filePath <- rownames(dataFiles)
} else {
  filePath <- paste0(gseDir, "/", list.files(path=gseDir))
}

# Load supplemental files as a data frame
gseCounts <- read.delim(filePath, header=TRUE)
```

## Cleaning the Data


## Normalizing the Data



## Mapping Gene Identifiers


## Clean-Up

```{r}
# Clear workspace
rm(list = ls())
```

## References

Isserlin, R. (2022). Lecture 3 - Finding Expression Data. Quercus.
https://q.utoronto.ca/courses/248455/files/18120910?module_item_id=3210846

My original R notebook

